---
title: "R Assignment 1"
author: "Azam && 18201501"
date: "12 October 2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Loading the dataset 
```{r}
EurostatCrime <- read.csv(file="EurostatCrime2015.csv", row.names = 1)
EurostatCrime
```

## Size and the structure of the dataset
```{r}
dim(EurostatCrime)
str(EurostatCrime)
```

## Add new column Sex.crime to the dataset
```{r}
EurostatCrime1 <- EurostatCrime
str(EurostatCrime1)
EurostatCrime1$Sex.crime <- EurostatCrime1$Rape + EurostatCrime1$Sexual.assault + EurostatCrime1$Sexual.violence
head(EurostatCrime1)
```

## Remove the colm Rape, Sexual.assault and Sexual.violence
```{r}
EurostatCrime1[,c("Rape", "Sexual.assault", "Sexual.violence")] <- list(NULL)
EurostatCrime1
```

## Countries with missing dataframe
```{r}
complete.cases(EurostatCrime1)
Countries_with_missing_dataset <- EurostatCrime1[!complete.cases(EurostatCrime1),]
row.names(Countries_with_missing_dataset)
```

## Removing countries with missing date from data frame
```{r}
DS_missing_data_removed_1 <- EurostatCrime1[complete.cases(EurostatCrime1),]
DS_missing_data_removed_1
DS_missing_data_removed_2 <- na.omit(EurostatCrime1)
DS_missing_data_removed_2
```

## Size of new data frame
```{r}
str(EurostatCrime1)
str(DS_missing_data_removed_1)
str(DS_missing_data_removed_2)
```

## Most common crime in Ireland in 2015
```{r}
Common_crime <- EurostatCrime1["Ireland",]
crime1 <- sort(Common_crime)
colnames(crime1[5])
```

## 3 Least common crime in Ireland in 2015
```{r}
crime2 <- sort(Common_crime)
for(i in seq(1,3)){
  print(colnames(crime2[i]))
}
```

## Country with maximum crime
```{r}
EurostatCrime1$Max.crime <- EurostatCrime1$Intentional.homicide + EurostatCrime1$Robbery + EurostatCrime1$Sex.crime + EurostatCrime1$Assault + EurostatCrime1$Theft
head(EurostatCrime1)
maximum <- which.max(EurostatCrime1$Max.crime)
row.names(EurostatCrime1)[maximum]
```

## # Simple Plot
The Plot reveals the Rape cases invloved in selected European countries during the year 2015.

```{r}
EurostatCrime <- read.csv(file="EurostatCrime2015.csv")
EurostatCrime2 <- EurostatCrime
boxplot(EurostatCrime2$X, EurostatCrime2$Rape, main="Boxplot")
x <- factor(EurostatCrime2$X)
y <- factor(EurostatCrime2$Rape)
plot(x,y, main="Plot of Rape cases in Europe with respect to countries")
```


## Pie Chart

The below Pie chart represents the pictorial representation of Sexual assaults happend in European countries
during the year 2015.

```{r}
slices <- c(40.36, 74.65, 34.99, 19.03, 4.52, 7.48)
lbls <- c("Austria", "Slovenia", "Bulgary", "Croatia", "Sweden", "Switzerland")
#pie(slices, labels = lbls, main="Sexual assault in European countries")
pct <- slices
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(slices,labels = lbls, col=rainbow(length(lbls)),
    main="Sexual assault in European countries in 2015")
```

## Bar chart

The Bar chart presented below picturizes the Sexual voilence took place in 2015 at few European countries chosen below

```{r}
findings <- factor(rep(c("Slovakia", "Spain", "Switzerland", "Austria"), c(11.90, 21.25, 32.90, 40.57)))
par(mar = c(3, 7, 1, 1)) # Adjust the margins on the four sides of the plot
barplot(height = table(findings), main = "Sexual Voilences in Ireland'15", xlab = "Number of stores",
        col = 1:4, horiz = TRUE, las = 1)
```