---
title: "R Assigment 2"
author: "Azam Jainullabudin Mohamed & 18201501"
date: "14 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.Load the lawyers' data into R. What proportion of the lawyers practices litigation
law? (Give your answer to 2 decimal places.)
```{r}
Lawyers <- read.csv("Lawyers.csv") #Importing the Lawyer.csv dataset
head(Lawyers)
Lawyers$Practice
Lawyers[,"Practice", drop=FALSE] #Extracting the Practice column
#Proportion of Lawyers practicing ligitation law
ligitation_law <- mean(Lawyers$Practice == "Litigation")
ligitation_law
#Proportion of Lawyers practicing ligitation law to 2 decimal places
proporation <- round(mean(Lawyers$Practice == "Litigation"), digits = 2) 
proporation
```

2. Is the proportion of lawyers in the Boston office that practice corporate law higher
than the proportion of lawyers in the Providence office that practice corporate law?
```{r}
# Lawyers practicing Corporate law at Boston office
Lawyers[,c("Office", "Practice"), drop=FALSE]
#subset(Lawyers, subset = Practice == 'Litigation')

# Lawyers practicing Corporate law at Boston office
boston_lawyers <- Lawyers[which(Lawyers$Office == 'Boston' & Lawyers$Practice == 'Corporate'),]
boston_lawyers

# Lawyers practicing Corporate law at Providence office
corporate_lawyers <- Lawyers[which(Lawyers$Office == 'Providence' & Lawyers$Practice == 'Corporate'),]
corporate_lawyers

#Count of Boston office lawyers
boston_length <- length(boston_lawyers$Office)
boston_length

#Count of Corporate office lawyers
corporate_length <- length(corporate_lawyers$Office)
corporate_length

# Sum of both Boston and Corporate office
total_proporation <- length(boston_lawyers$Office) +length(corporate_lawyers$Office)
total_proporation

#Proporation of Boston office
boston_proportion <- boston_length/total_proporation
boston_proportion

#Proporation of Corporate office
corporate_length <- corporate_length/total_proporation
corporate_length

#Verifying if Boston office proporation is greater than Corporate office proportion
boston_proportion > corporate_length
```

3.Use the aggregate function to compute the average age of lawyers who practice
corporate law and of lawyers who practice litigation law, across the different levels
of seniority. Label the columns of the resulting data frame appropriately.
```{r}
# average age of lawyers practising corporate law and litigation law, accross seniority
avg_age <- Lawyers[which(Lawyers$Practice == 'Litigation' | Lawyers$Practice == 'Corporate'),]

# Using Aggregate function to manipulate the mean of the above match
output_mean <- aggregate(avg_age$Age, by=list(avg_age$Practice, avg_age$Seniority), FUN =mean)

# Renaming the column names of Group1 and Group2
colnames(output_mean)[1] <- "Law Practice"
colnames(output_mean)[2] <- "Seniority level"
output_mean
```

4. Which office has the youngest median age?
```{r}
# Extracting the lawyers median age
median_age <- median(Lawyers$Age)
median_age

# Office with young median age
young <- Lawyers[which(Lawyers$Age == median_age), "Office"]
young
```

Task 2: Writing your own function
1. Write a function which compute the Rosenbrock banana function using a loop. Test
the function on the vectors x = (.2, .5) and x = (.2, .5, .1, .6)

```{r}
# Function for manipulating Rosenbrock banana function using loop
rosenbrock <- function(x){
  # Updating the summation end limit as n-1
  n <- length(x)-1
  z <- 0
  s <- 0
  sum <- 0
  for(i in 1:n){
    # Expression for rosenbrock banana function
    s = (100 * (x[i+1] - x[i] * x[i])^2 + (1 - x[i])^2)
    z[i] <- s
  }
  for(j in 1:n){
    # Finding the summation
    sum = sum + z[j]
    print(sum)
  }
}
# Tesing on Vector x
x <- c(0.2,0.5)
print("Rosenbrock function with vector x")
rosenbrock(x)

# Tesing on vector y
y <- c(0.2,0.5,0.1,0.6)
print("Rosenbrock function with vector y")
rosenbrock(y)
```

2. Propose an alternative function that does not use any loop. Test the function on
the same two vectors.

```{r}
# Function for manipulating Rosenbrock banana function without using loop
sum <- 0
banana_function <- function(x){
  n <- length(x)-1 # Updating the summation end limit as n-1
  i <- 1:n
  m = 100 * (x[i+1] - x[i] * x[i])^2 + (1 - x[i])^2
  # Cummulative sum for calcating the summation values
  cumsum(m)
}
# Tesing on Vector x
x <- c(0.2,0.5)
print("Rosenbrock function with vector x")
banana_function(x)

y <- c(0.2,0.5,0.1,0.6)
print("Rosenbrock function with vector y")
banana_function(y)
```

3.Compare the timings you obtain by repeating the function calls 100 times using the
vector x = (.2, .5, .1, .6) as input.
```{r}
# Printing the timings obtained
system.time(replicate(100, rosenbrock(x = c(0.2,0.5,0.1,0.6))))
system.time(replicate(100, banana_function(x = c(0.2,0.5,0.1,0.6))))
```

Task 3: Writing S3 methods
1. Load in the data as an object called DublinAirport. Assign to the DublinAiport
object the classes WeatherData and data.frame.

```{r}
#Importing the DublinAirport.csv dataset
DublinAirport <- read.csv("2018_09_Dublin_Airport.csv")
# Assigning the data frame DublinAirport to the class
class(DublinAirport) <- c('WeatherData', 'data.frame')
DublinAirport
```

2. Write an S3 summary method for an object of class WeatherData which produces
the following statistical summaries for the rain, maxtp, mintp variables: mean,
standard deviation, minimum, maximum.
```{r}
# Writing a new summary method for class WeatherData
summary.WeatherData <- function(object){
  # Calculating the mean, sd, max, min value of "Rain" using sapply and
  # printing the result using cat command
  cat('Mean of Rain', sapply(object["rain"], mean), '\n')
  cat('Standart deviation of Rain', sapply(object["rain"], sd), '\n')
  cat('Max of Rain', sapply(object["rain"], max), '\n')
  cat('Min of Rain', sapply(object["rain"], min), '\n')
  
  # Calculating the mean, sd, max, min value of "maxtp" using sapply and
  # printing the result using cat command
  cat('Mean of maxtp', sapply(object["maxtp"], mean), '\n')
  cat('Standart deviation of Maxtp', sapply(object["maxtp"], sd), '\n')
  cat('Max of Maxtp', sapply(object["maxtp"], max), '\n')
  cat('Min of Maxtp', sapply(object["maxtp"], min), '\n')

  # Calculating the mean, sd, max, min value of "mintp" using sapply and
  # printing the result using cat command
  cat('Mean of Mintp', sapply(object["mintp"], mean), '\n')
  cat('Standart deviation of Mintp', sapply(object["mintp"], sd), '\n')
  cat('Max of Mintp', sapply(object["mintp"], max), '\n')
  cat('Min of Mintp', sapply(object["mintp"], min), '\n')
}
summary.WeatherData(DublinAirport)
```

3. Download the new data set 2018 09 Cork Airport.csv from Blackboard, assign
the classes WeatherData and data.frame to the object containing the Cork data,
and test your function on it. Interpret your findings for Dublin and Cork Airports.
```{r}
#Importing the DublinAirport.csv dataset
CorkAirport <- read.csv("2018_09_Cork_Airport.csv")
# Assigning the data frame DublinAirport to the class
class(CorkAirport) <- c('WeatherData', 'data.frame')

summary.WeatherData(CorkAirport)
```
Interpreting both Dublin and Cork dataset, its evident that CorkAirport overweighs that DublinAirport with respect to the precipitation amount based on mean, standard deviationg, minimum, maximum and viceversa with respect to Maximum Air Temperature.
For minimum air temperature, standart deviation and maximum of DublinAirport is greater than CorkAirport dataset, on the other hand,
mean and minimum of CorkAirport leads DublinAirport dataset.

4. Create an S3 plot method for the class WeatherData that produces the following plots.
```{r}
plot.WeatherData <- function(object, user, color1, color2){
  # For printing the plot in single panel
  par(mfrow=c(2,1))
  if(all(object == DublinAirport) & missing(user)){
    #For ploting the axis with limits
    plot(object$ï..date, xlab = "Dates", ylab = "Maximum Air Temperature", main = "Dublin WeatherData", ylim=c(0,25))
  }
  else if(all(object == CorkAirport) & missing(user)){
    #For ploting the axis with limits
    plot(object$ï..date, xlab = "Dates", ylab = "Maximum Air Temperature", main = "Cork WeatherData", ylim=c(0,25))
  }
  else if(all(object == DublinAirport) & user==TRUE){
    #For ploting the axis with limits
    plot(object$ï..date, xlab = "Dates", ylab = "Maximum Air Temperature", main = "Dublin WeatherData (User Modified)", ylim=c(0,25))
  }
  else if(all(object == CorkAirport) & user==TRUE){
    #For ploting the axis with limits
    plot(object$ï..date, xlab = "Dates", ylab = "Maximum Air Temperature", main = "Cork WeatherData (User Modified)", ylim=c(0,25))
  }
  #For generating the Maximum temperature plot using red line
  lines(object$maxtp, col="red")
  #For generating the Minimum temperature plot using blue line
  lines(object$mintp, col="blue")
  
  #To mention the legend of the plot
  legend("topright", legend=c("Max.Temperature", "Min.Temperature"), col=c("red", "blue"), lty=1, cex=0.8)
  # For printing the vertical line
  abline(v=object$ï..date, lty=2, col="grey")
  #Hightlighting the highest maximum temperature registered
  points(object[which.max(object$maxtp),]$ï..date,object[which.max(object$maxtp),]$maxtp, pch=19, bg="black",col="black", lwd=10)
  #Hightlighting the highest maximum temperature registered using vertical line
  abline(v=object[which.max(object$maxtp),]$ï..date ,col='darkorange1', lwd=2)
  #Hightlighting the lowest minimum temperature registered
  points(object[which.min(object$mintp),]$ï..date,object[which.min(object$mintp),]$mintp, pch=19, bg="black", col="black", lwd=10)
  #Hightlighting the lowest minimum temperature registered using vertical line
  abline(v=object[which.min(object$mintp),]$ï..date ,col="darkorange1", lwd=2)

  # USer control to change the color
  if(!missing(user)){
    #For generating the Maximum temperature plot using user input color1
    lines(object$maxtp, col=color1)
    #For generating the Minimum temperature plot using user input color2
    lines(object$mintp, col=color2)
    #To mention the legend of the plot as per user's input
    legend("topright", legend=c("Max.Temperature", "Min.Temperature"), col=c(color1, color2), lty=1, cex=0.8)
  }
  
  #Bottom plot
  # Ploting the Precipation plot with vertical line
  plot(as.numeric(object$ï..date), object$rain, xlab="Date", ylab="Precipitation", type="h", lwd=10)
  # Highlighting the highest amount of rainfall in red
  points(as.numeric(object[which.max(object$rain),]$ï..date),object[which.max(object$rain),]$rain, type="h", bg="red",col="red", lwd=10)

  # USer control to change the color
  if(!missing(user)){
    points(as.numeric(object[which.max(object$rain),]$ï..date),object[which.max(object$rain),]$rain, type="h", col=color1, lwd=10)
  }  
}
plot.WeatherData(DublinAirport, )
plot.WeatherData(DublinAirport, user=TRUE, color1 = "green", color2= "dark red")
plot.WeatherData(CorkAirport)
plot.WeatherData(CorkAirport, user=TRUE, color1 = "yellow", color2= "green")
```
